---
title: "P8105_hw2_hc3212"
author: "Hening cui"
date: "2021/10/3"
output: github_document
---

```{r library, echo = FALSE, include=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

## Problem 1

Read and clean the Mr. Trash Wheel sheet.

```{r trash_wheel_df}
trash_wheel_df = 
  read_excel("./data/Trashwheel.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

Read and clean precipitation data for 2018 and 2019.

```{r pre_2018_2019}
pre_2018_df = 
  read_excel("./data/Trashwheel.xlsx", sheet = "2018 Precipitation", range = "A2:B15") %>%
  janitor::clean_names() %>% 
  drop_na(month, total) %>% 
  mutate(year = 2018) %>% 
  relocate(year) 

pre_2019_df = 
  read_excel("./data/Trashwheel.xlsx", sheet = "2019 Precipitation", range = "A2:B15") %>%
  janitor::clean_names() %>% 
  drop_na(month, total) %>% 
  mutate(year = 2019) %>%
  relocate(year) 
```

Precipitation combine.

```{r merging_pre}
NPre_df= 
  bind_rows(pre_2018_df, pre_2019_df) %>% 
  mutate(month = month.name[month])
```

The first data set was generated by Mr. Trashwheel trash collector in Baltimore, Maryland which include the information of trash entering, colletion and dumpster storage. The dataset contains information on `r colnames(trash_wheel_df)`. There are a total of `r nrow(trash_wheel_df)` rows in our final dataset. 

The second data set was the month precipitation data in 2018 and 2019. The dataset contains information on `r colnames(NPre_df)`. There are a total of `r nrow(NPre_df)` rows in our final dataset.


* The total precipitation in 2018 was `r NPre_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.
* The median number of sports balls found in a dumpster in 2019 was `r trash_wheel_df %>% filter(year == c("2019")) %>% pull(sports_balls) %>% median()`

## Problem 2

Read and clean pols-month.csv.

```{r pols_month_df, message = FALSE}
pols_month_df = 
  read_csv("./data/fivethirtyeight/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month = month.name[as.integer(month)]) 

pols_month_clean_df = 
  pivot_longer(
    pols_month_df, 
    prez_gop:rep_dem,
    names_to  = "president",
    values_to  = "values"
  ) %>% 
  filter(!president %in% c("prez_dem","prez_gop")) %>% 
  select(-day)
view(pols_month_clean_df)
```

Read and clean snp.csv.

```{r snp_df, message = FALSE}
snp_df = 
  read_csv("./data/fivethirtyeight/snp.csv") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(date = mdy(as.character(date))) %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month = month.name[as.integer(month)]) %>% 
  select(-day) %>% 
  arrange(year, month) %>% 
  relocate(year, month)
view(snp_df) 
```

Read and clean unemployment.csv.

```{r unemployment_df,message = FALSE}
unemployment_df = 
  read_csv("./data/fivethirtyeight/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  drop_na() 

unemployment_clean_df =
  pivot_longer(
    unemployment_df,
    jan:dec,
    names_to = "month",
    values_to = "percentage"
  ) 
view(unemployment_clean_df)
  
```

Join the datasets.

```{r merging_Fivethirtyeight, message = FALSE}
pols_tidy_df =
  pols_month_clean_df %>% 
  mutate(month = paste(month, "01", sep = "-"),
         month = as.Date(month, format = "%B-%d"),
         month = month(month)
         ) %>% 
  unite(date,year,month,sep = "-")
view(pols_tidy_df) 

snp_tidy_df =
  snp_df %>% 
  mutate(month = paste(month, "15", sep = "-"),
         month = as.Date(month, format = "%B-%d"),
         month = month(month)
         ) %>% 
  unite(date,year,month,sep = "-")
view(snp_tidy_df) 

unemployment_tidy_df =
  snp_df %>% 
  mutate(month = paste(month, "01", sep = "-"),
         month = as.Date(month, format = "%B-%d"),
         month = month(month)
         ) %>% 
  unite(date,year,month,sep = "-")
view(unemployment_tidy_df) 

pols_snp_df = 
  full_join(pols_tidy_df,snp_tidy_df,by = c("date" = "date")) 
view(pols_snp_df)

Fivethirtyeight_df =
  left_join(pols_snp_df,unemployment_tidy_df, by = c("date" = "date"))
Fivethirtyeight_df
```

The pols-month dataset describes the political landscape information of the US from `r min(pull(pols_month_clean_df, year))` to `r max(pull(pols_month_clean_df, year))`. There are a total of `r nrow(pols_tidy_df)` rows, and `r ncol(pols_tidy_df)` columns in our final dataset. The dataset contains information on `r colnames(pols_tidy_df)`.

The snp dataset shows the closing vales of the S&P stock index on the associated month and year which from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`. There are a total of `r nrow(snp_tidy_df)` rows, and `r ncol(snp_tidy_df)` columns in our final dataset. The dataset contains information on `r colnames(snp_tidy_df)`.

The unemployment dataset presents the unemployment percentage of a particular month  from `r min(pull(unemployment_clean_df, year))` to `r max(pull(unemployment_clean_df, year))`. There are a total of `r nrow(unemployment_clean_df)` rows, and `r ncol(unemployment_clean_df)` columns in our final dataset. The dataset contains information on `r colnames(unemployment_clean_df)`.

The Fivethirtyeight dataset contains information on the political landscape, stock performance, and unemployment percentages by merging the previos 3 datasets together according to date. The time range from `r min(substring(pull(Fivethirtyeight_df,date), 1, 4))` to `r max(substring(pull(Fivethirtyeight_df,date), 1, 4))`. There are a total of `r nrow(Fivethirtyeight_df)` rows, and `r ncol(Fivethirtyeight_df)` columns in our final dataset. The dataset contains information on `r colnames(Fivethirtyeight_df)`.

## Problem 3

Read and clean the NYC Popular_Birth_Name.csv.

```{r NYC_birthname_df, message = FALSE}
NYC_birthname_df = 
  read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%  
  drop_na() %>%
  mutate(childs_first_name = toupper(childs_first_name), 
         ethnicity = gsub("BLACK NON HISP$", "BLACK NON HISPANIC", pull(.,ethnicity)),
         ethnicity = gsub("WHITE NON HISP$", "WHITE NON HISPANIC", pull(.,ethnicity)),
         ethnicity = gsub("ASIAN AND PACI$", "ASIAN AND PACIFIC ISLANDER", pull(.,ethnicity))) %>% 
  distinct()
view(NYC_birthname_df)
```

Build Olivial popularity table.

```{r Olivial, message = FALSE}
Olivial_df = 
  filter(NYC_birthname_df, childs_first_name == "OLIVIA", gender =="FEMALE") %>% 
  select(-count) %>% 
  knitr::kable()

```



Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. Produce a similar table showing the most popular name among male children over time.

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).



