---
title: "P8105_hw2_hc3212"
author: "Hening cui"
date: "2021/10/3"
output: github_document
---

```{r library, echo = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Read and clean the Mr. Trash Wheel sheet.

```{r}
trash_wheel_df = 
  read_excel("./Trashwheel.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(sports_balls))
view(trash_wheel_df)
```

Read and clean precipitation data for 2018 and 2019.

```{r}
pre_2018_df = 
  read_excel("./Trashwheel.xlsx", sheet = "2018 Precipitation", range = "A2:B15") %>%
  janitor::clean_names() %>% 
  drop_na(month, total) %>% 
  mutate(year = 2018) %>% 
  relocate(year) 
view(pre_2018_df)

pre_2019_df = 
  read_excel("./Trashwheel.xlsx", sheet = "2019 Precipitation", range = "A2:B15") %>%
  janitor::clean_names() %>% 
  drop_na(month, total) %>% 
  mutate(year = 2019) %>%
  relocate(year) 
view(pre_2019_df)
```

Precipitation combine.

```{r}
NPre_df= 
  bind_rows(pre_2018_df, pre_2019_df) %>% 
  mutate(month = month.name[month])
```

The first data set was generated by Mr. Trashwheel trash collector in Baltimore, Maryland which include the information of trash entering, colletion and dumpster storage. The dataset contains information on `r colnames(trash_wheel_df)`. There are a total of `r nrow(trash_wheel_df)` rows in our final dataset. 

The second data set was the month precipitation data in 2018 and 2019. The dataset contains information on `r colnames(NPre_df)`. There are a total of `r nrow(NPre_df)` rows in our final dataset.


* The total precipitation in 2018 was `r NPre_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.
* The median number of sports balls found in a dumpster in 2019 was `r trash_wheel_df %>% filter(year == c("2019")) %>% pull(sports_balls) %>% median()`





